---
title: "002-PP-BS-dil_series"
format: html
identifier: "002"
filters:
   - lightbox
lightbox: auto
---

# Short Description

This is a the batch culture well plate experiment, with serial dilutions of starting densities of *B. subtilis* and *P. putida* monocultures. Starting densities ranged from 10^-1^ to 10^-6^ and 10^-2^ to 10^-7^ OD~600~, respectively. Culturing was carried out on plate reader with continuous measurements (every 1h) for \~24 h.

# Data loading and `df` arrangement

```{r}
#| output: false

library(tidyverse)
library(readxl)
library(data.table)
library(patchwork)
library(stringr)
library(RColorBrewer)
library(ggnested)
library(growthcurver)
library(ggtext)
```

```{r}
#| output: false

# Step 1: Load your Excel file
OD <- read_excel("002-PP_BS_dil_series_24h_031826.xlsx", range = "B122:CU147", col_names = TRUE)  
G <- read_excel("002-PP_BS_dil_series_24h_031826.xlsx", range = "B178:CU203", col_names = TRUE) # GFP70
R <- read_excel("002-PP_BS_dil_series_24h_031826.xlsx", range = "B206:CU231", col_names = TRUE) # RFP90

# Step 2: Reshape from wide to long format
OD <- OD[-2]                 # Remove Temp column
OD_long <- pivot_longer(
  data = OD,
  cols = -1,                 # All columns except the first (time)
  names_to = "well",
  values_to = "OD600"
)

G <- G[-2]                 # Remove Temp column
G_long <- pivot_longer(
  data = G,
  cols = -1,                 # All columns except the first (time)
  names_to = "well",
  values_to = "G"
)

R <- R[-2]                 # Remove Temp column
R_long <- pivot_longer(
  data = R,
  cols = -1,                 # All columns except the first (time)
  names_to = "well",
  values_to = "R"
)

# Step 3: Rename first column to "time" 
colnames(OD_long)[1] <- "time"
colnames(G_long)[1] <- "time"
colnames(R_long)[1] <- "time"

# Step 4 (optional): Convert time to numeric and to hours
OD_long$time <- as.numeric(OD_long$time)
G_long$time <- as.numeric(G_long$time)
R_long$time <- as.numeric(R_long$time)

OD_long$time <- OD_long$time/3600
G_long$time <- G_long$time/3600
R_long$time <- R_long$time/3600

OD_long$G <- G_long$G
OD_long$R <- R_long$R

OD_long <- OD_long %>%
  mutate(
    column = as.numeric(str_extract(well, "\\d+"))
  )

OD_long <- OD_long %>%
  mutate(
    dilution = case_when(
      column %in% 1  ~ "10^-1",
      column %in% 2  ~ "10^-2",
      column %in% 3  ~ "10^-3",
      column %in% 4  ~ "10^-4",
      column %in% 5  ~ "10^-5",
      column %in% 6  ~ "10^-6",
      column %in% 7  ~ "10^-2",
      column %in% 8  ~ "10^-3",
      column %in% 9  ~ "10^-4",
      column %in% 10  ~ "10^-5",
      column %in% 11  ~ "10^-6",
      column %in% 12  ~ "10^-7"
    )
  )

OD_long <- OD_long %>%
  mutate(
    species = case_when(
      column %in% 1:6  ~ "Pp",
      column %in% 7:12  ~ "Bs"
    )
  )


OD_long <- OD_long %>%
  mutate(condition = paste(species, dilution, sep = " "))

# Make dilution a properly ordered factor
OD_long$dilution <- factor(OD_long$dilution, levels = sort(unique(OD_long$dilution)))
```

## Summary stats

```{r}
summary_df <- OD_long %>%
  group_by(time, species, dilution) %>%
  summarise(
    mean_OD = mean(OD600, na.rm = TRUE),
    sd_OD   = sd(OD600, na.rm = TRUE),
    n       = n(),
    se_OD   = sd_OD / sqrt(n),
    mean_G = mean(G, na.rm = TRUE),
    sd_G   = sd(G, na.rm = TRUE),
    se_G   = sd_G / sqrt(n),
    mean_R = mean(R, na.rm = TRUE),
    sd_R   = sd(R, na.rm = TRUE),
    se_R   = sd_R / sqrt(n),
    .groups = "drop"
  )

# Make dilution a properly ordered factor (AGAIN)
summary_df$dilution <- factor(summary_df$dilution, levels = sort(unique(summary_df$dilution)))
```

## Write/Load CSV

```{r}
write.csv(summary_df, "summary_df.csv", row.names = FALSE)
write.csv(OD_long, "df.csv", row.names = FALSE)

```

**Load df and summary_df from CSV:**

```{r}
df <- read.csv ("df.csv") 
summary_df <- read.csv ("summary_df.csv") 
```

# Data Visualization

## OD~600~

### OD all, facetted

```{r}
#| label: 002-PPBS_allOD.facet

ggplot(summary_df, aes( x = time, y = mean_OD, color = dilution, linetype = species )) + 
  geom_line() +
  geom_point() + 
  geom_errorbar( aes(ymin = mean_OD - se_OD, ymax = mean_OD + se_OD), width = 0.2 ) +
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme_minimal() + 
  labs(
  y=expression("OD"[600]), 
  x="Time [h]", 
  title = "OD<sub>600</sub> of *P. putida* and *B. subtilis* for a serial dilution of starting densities") + 
  facet_wrap(~ species, scales = "free_y") +
  theme(plot.title = element_markdown())
```

### OD selected dilutions

```{r}
#| label: 002-PPBS_selectedDilutions

summary_df %>% filter( (dilution %in% c("10^-6") & species %in% c("Pp")) | (dilution %in% c("10^-2") &
species %in% c("Bs")) ) %>% 
  ggplot(aes(x = time, y = mean_OD, color = dilution, linetype = species, group = interaction(species, dilution))) +
  geom_line() + 
  geom_point() + 
  geom_errorbar(aes(ymin = mean_OD - se_OD,
ymax = mean_OD + se_OD), width = 0.2 ) + 
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme_minimal() + labs(
    y= expression("OD"[600]), 
    x="Time [h]", 
    title = "OD<sub>600</sub> of *P. putida* and *B. subtilis* comparison for the biggest diference in starting densities") +
  theme(plot.title = element_markdown())
```

### OD all, with ggnested

```{r}
#| label: 002-allOD.ggnested

ggnested(summary_df, aes( x = time, y = mean_OD, main_group = species, sub_group = dilution ),
         #legend_labeling = "join", #join_str = ": ",
         legend_title = "Species: Starting Frequencies", main_keys = F, base_clr = "green", gradient_type = "shades" ) + 
  geom_line() + 
  geom_point() +
  geom_errorbar( aes(ymin = mean_OD - se_OD, ymax = mean_OD + se_OD), width = 0.2 ) +
  #scale_colour_discrete(labels = scales::label_parse()) +
  # scale_color_discrete(labels = function(x) parse(text = x)) +
  theme_minimal() + 
  labs(y= expression("OD"[600]), x="Time [h]", 
       title = expression("OD"[600]*" of Pp and Bs for different starting frequencies")) 
```

## PP Plot calculations for different wavelength

```{r}
PPOD <- summary_df %>% filter(species == "Pp") %>% 
  ggplot(aes(x = time, y = mean_OD, color = dilution)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = mean_OD - se_OD, ymax = mean_OD + se_OD), width = 0.2 ) + 
  theme_minimal() +
  labs(y= expression("OD"[600]), x="Time [h]") + 
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme(plot.title = element_markdown())

PPG <- summary_df %>% filter(species == "Pp") %>% 
  ggplot(aes(x = time, y = mean_G, color = dilution)) +
  geom_line() +
  geom_point() +
  geom_errorbar( aes(ymin = mean_G - se_G, ymax = mean_G + se_G), width = 0.2 ) +
  theme_minimal() + theme(legend.position = 'none') +
  labs(y= "GFP", x="Time [h]") + 
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme(plot.title = element_markdown())

BSR <- summary_df %>% filter(species == "Pp") %>% 
  ggplot(aes(x = time, y = mean_R, color = dilution)) + 
  geom_line() + 
  geom_point() +
  geom_errorbar( aes(ymin = mean_R - se_R, ymax = mean_R + se_R), width = 0.2 ) + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(y= "RFP", x="Time [h]") + 
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme(plot.title = element_markdown())
```

## BS Plot calculations for different wavelength

```{r}
BSOD <- summary_df %>% filter(species == "Bs") %>% 
  ggplot(aes(x = time, y = mean_OD, color = dilution)) + 
  geom_line() + 
  geom_point() +
  geom_errorbar(aes(ymin = mean_OD - se_OD, ymax = mean_OD + se_OD), width = 0.2 ) + 
  theme_minimal() +
  labs(y= expression("OD"[600]), x="Time [h]") + 
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme(plot.title = element_markdown())

BSR <- summary_df %>% filter(species == "Bs") %>% 
  ggplot(aes(x = time, y = mean_R, color = dilution)) + 
  geom_line() + 
  geom_point() +
  geom_errorbar( aes(ymin = mean_R - se_R, ymax = mean_R + se_R), width = 0.2 ) + 
  theme_minimal() + 
  theme(legend.position = "none") + 
  labs(y= "RFP", x="Time [h]") + 
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme(plot.title = element_markdown())

BSG <- summary_df %>% filter(species == "Bs") %>% 
  ggplot(aes(x = time, y = mean_G, color = dilution)) + 
  geom_line() + 
  geom_point() + 
  theme_minimal() + 
  geom_errorbar( aes(ymin = mean_G - se_G, ymax = mean_G + se_G), width = 0.2 ) + 
  theme(legend.position = "none") + labs(y= "GFP", x="Time [h]") + 
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  theme(plot.title = element_markdown())
```

## *P. putida* OD & GFP

```{r}
#| label: 002-PP_OD_GFP.facet

PPOD + PPG + 
  plot_annotation( title = "Growth of *P. putida* by OD<sub>600</sub> turbidity (A) and GFP intensity (B) for different inoculation frequencies", tag_levels = "A") + 
  plot_layout(guides = 'collect') & 
  theme( plot.title = element_markdown(), legend.position = 'bottom') & labs(color = expression("OD"[600]*" at inoculation"))
```

### *B. subtilis* OD & RFP

```{r}
#| label: 002-BS_OD_RFP.facet

BSOD + BSR + 
  plot_annotation( title = "Growth of *B. subtilis* by OD<sub>600</sub> turbidity (A) and RFP intensity (B) for different inoculation frequencies", tag_levels = "A") + 
  plot_layout(guides = 'collect') & 
  theme( plot.title = element_markdown(), legend.position = 'bottom') & 
  guides(color = guide_legend(nrow = 1)) & 
  labs(color = expression("OD"[600]*" at inoculation"))
```

### *B. subtilis* OD, RFP & GFP

```{r}
#| label: 002-BS_OF_RFP_GFP.facet

BSOD + BSR + BSG + 
  plot_annotation(
  title = "Growth of *B. subtilis* by OD<sub>600</sub> turbidity (A), RFP (B) and GFP (C) intensity for different inoculation frequencies", tag_levels = "A") +
  plot_layout(guides = 'collect') & 
  theme( plot.title = element_markdown(), legend.position = 'bottom' )
```

## Fluorescence per OD

Calculate fluorescence per OD:

```{r}
summary_df$`GFP/OD` <- summary_df$mean_G / summary_df$mean_OD
summary_df$`RFP/OD` <- summary_df$mean_R / summary_df$mean_OD
```

### *P. putida* GFP/OD with growth phase indicators

```{r}
#|label: 002-PP_GFP_per_OD

summary_df %>% filter(species == "Pp") %>% 
  ggplot(aes(x=time, y=`GFP/OD`, colour = dilution)) + 
  geom_point() + 
  geom_line() +
  theme_minimal() + 
  theme(plot.title = element_markdown()) +
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  labs( title = "GFP fluorescence per cell density (OD<sub>600</sub>) of a *P. putida* serial dilution", subtitle = "Growth phase indicators according to 10^-1 dilution", 
        x = "Time [h]", y = expression("GFP"/"OD"[600]), 
        color = expression("Initial OD"[600]) ) + 
  
  # Growth phase indicators
  geom_vline(xintercept = c(4.5, 11), linetype = "dashed", color = "#8C8C8C") + annotate("text", x = 2.25, y = min(summary_df$`GFP/OD`) * 1.1, label = expression(lag), color = "#8C8C8C") +
  annotate("text", x = 8, y = min(summary_df$`GFP/OD`) * 1.1, label = expression(log), color = "#8C8C8C") + 
  annotate("text", x = 16, y = min(summary_df$`GFP/OD`) * 1.1, label = "stationary", color = "#8C8C8C")
```

### *P. putida* RFP/OD with growth phase indicators

```{r}
#| label: 002-PP_RFP_per_OD

summary_df %>% filter(species == "Pp") %>% 
  ggplot(aes(x=time, y=`RFP/OD`, colour = dilution)) + 
  geom_point() + 
  geom_line() +
  theme_minimal() + 
  theme(plot.title = element_markdown()) +
  scale_color_discrete(labels = function(x) parse(text = x)) + 
  labs(
    title = "RFP fluorescence per cell density (OD<sub>600</sub>) of a *P. putida* serial dilution",
    subtitle = "Growth phase indicators according to 10^-1 dilution", 
    x = "Time [h]", y = expression("RFP"/"OD"[600]), color = expression("Initial OD"[600]) ) + 
  
  # Growth phase indicators
  geom_vline(xintercept = c(4.5, 11), linetype = "dashed", color = "#8C8C8C") + 
  annotate("text", x = 2.25, y = min(summary_df$`GFP/OD`) * 1.1, label = expression(lag), color = "#8C8C8C") +
  annotate("text", x = 8, y = min(summary_df$`GFP/OD`) * 1.1, label = expression(log), color = "#8C8C8C") + 
  annotate("text", x = 16, y = min(summary_df$`GFP/OD`) * 1.1, label = "stationary", color = "#8C8C8C")
```

# Growth Rates

```{r}
fit_gc <- function(df) {
  tryCatch({
    growthcurver::SummarizeGrowth(
      data_t = as.numeric(df$time),
      data_n = as.numeric(df$mean_OD),
      bg_correct = "min"
    )
  }, error = function(e) {
    message("Failed fit")
    return(NULL)
  })
}

gc_fits <- summary_df %>%
  group_by(species, dilution) %>%
  nest() %>%
  mutate(gc = map(data, fit_gc))
```

## Apply per group

```{r}
gc_params <- summary_df %>%
  group_by(species, dilution) %>%
  nest() %>%
  mutate(
    gc = map(data, fit_gc),
    t_mid = map_dbl(gc, ~ ifelse(is.null(.x), NA, .x$vals$t_mid)),
    r     = map_dbl(gc, ~ ifelse(is.null(.x), NA, .x$vals$r)),
    k     = map_dbl(gc, ~ ifelse(is.null(.x), NA, .x$vals$k))
  ) %>%
  ungroup()


bg_df <- summary_df %>%
  group_by(species, dilution) %>%
  summarise(bg = min(mean_OD, na.rm = TRUE), .groups = "drop")

gc_params <- gc_params %>%
  left_join(bg_df, by = c("species", "dilution"))
```

## Generate smooth fitted curves

```{r}
predict_growth <- function(species_name) {
  gc_params %>%
    filter(species == species_name) %>%
    filter(!is.na(k), !is.na(r), !is.na(t_mid)) %>%
    rowwise() %>%
    mutate(
      pred = list({
        time_seq <- seq(0, max(summary_df$time), length.out = 100)
        pred_vals <- k / (1 + exp(-r * (time_seq - t_mid)))
        tibble(time = time_seq, pred_OD = pred_vals + bg)
      })
    ) %>%
    ungroup() %>%
    select(dilution, pred) %>%
    unnest(pred)
}

PP_pred <- predict_growth("Pp")
BS_pred <- predict_growth("Bs")
```

## Plot: raw data + fitted curves

### PP

```{r}
#| output: false

PPfit <- summary_df %>%
  filter(species == "Pp") %>%
  ggplot(aes(x = time, y = mean_OD, color = dilution)) +
  geom_line(alpha = 0.4) +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_OD - se_OD, ymax = mean_OD + se_OD),
    width = 0.2
  ) +
  geom_line(
    data = PP_pred,
    aes(x = time, y = pred_OD, color = dilution),
    # linetype = "dashed",
    linewidth = 0.9
  ) +
  scale_color_discrete(labels = parse(text = unique(summary_df$dilution))) +
  theme_minimal() +
  labs(y = expression("OD"[600]), 
       x = "Time [h]",
       title = "Logistic growth curve fit"
       ) +
  theme(plot.title = element_markdown()) +
  scale_color_discrete(labels = function(x) parse(text = x))
```

### BS

```{r}
#| output: false

BSfit <- summary_df %>%
  filter(species == "Bs") %>%
  ggplot(aes(x = time, y = mean_OD, color = dilution)) +
  geom_line(alpha = 0.4) +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_OD - se_OD, ymax = mean_OD + se_OD),
    width = 0.2
  ) +
  geom_line(
    data = BS_pred,
    aes(x = time, y = pred_OD, color = dilution),
    # linetype = "dashed",
    linewidth = 0.9
  ) +
  scale_color_discrete(labels = parse(text = unique(summary_df$dilution))) +
  theme_minimal() +
  labs(y = expression("OD"[600]), 
       x = "Time [h]",
       title = "Logistic growth curve fit"
  ) +
  theme(plot.title = element_markdown()) +
  scale_color_discrete(labels = function(x) parse(text = x)) 
```

## Extract time of max growth rate (t₀)

```{r}
analysis_df <- gc_params %>%
  select(species, dilution, t_mid) %>%
  rename(t0 = t_mid)   # optional: keep old naming if you like

t0_points <- gc_params %>%
  filter(species == "Pp") %>%
  filter(!is.na(k), !is.na(r), !is.na(t_mid)) %>%
  rowwise() %>%
  mutate(
    OD_t0 = list({
      # Predicted OD at t_mid
      pred_OD <- k / (1 + exp(-r * (t_mid - t_mid))) + bg
      pred_OD
    }),
    start_OD_theoretical = 10^as.numeric(gsub("10\\^", "", dilution))
  ) %>%
  ungroup() %>%
  select(dilution, t_mid, OD_t0, start_OD_theoretical)

# Get starting OD per group
start_od <- summary_df %>%
  filter(time == 0) %>%
  select(species, dilution, start_OD = mean_OD)

# Combine
analysis_df <- analysis_df %>%
  left_join(start_od, by = c("species", "dilution"))

## Add theoretical calculated starting OD
analysis_df <- analysis_df %>%
  mutate(
    start_OD_theoretical = 10^as.numeric(gsub("10\\^", "", dilution))
  )
```

## Calculate plots and linear regression

### PP

```{r}
PPtime <- analysis_df %>%
  filter(species == "Pp") %>%
  ggplot(aes(x = start_OD_theoretical, y = t0, color = species)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_x_log10() +  # optional, since dilutions span orders of magnitude
  labs(
    y = "Time [h] until max. growth rate",
    x = expression("Theoretical Starting OD"[600]),
    # title = "*P. putida*"
    title = "Growth timing for different starting density"
  ) +
  theme_minimal() +
  theme(legend.position = "none", 
        plot.title = element_markdown()
        ) +
  scale_color_discrete(labels = function(x) parse(text = x))
```

### BS

```{r}
BStime <- analysis_df %>%
  filter(species == "Bs", dilution != "10^-7") %>%
  ggplot(aes(x = t0, y = start_OD_theoretical, color = species)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = TRUE) +
  scale_y_log10() +  # optional, since dilutions span orders of magnitude
  labs(
    x = "Time [h] until max. growth rate",
    y = expression("Theoretical Starting OD"[600]),
    # title = "*B. subtilis*"
    title = "Growth timing for different starting density"
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_markdown()
        ) +
  scale_color_discrete(labels = function(x) parse(text = x))
```

Fit regression explicitly:

```{r}
lm_model <- lm(t0 ~ start_OD_theoretical, data = analysis_df %>% filter(species == "Pp"))
summary(lm_model)
```

## Plot with patchwork

```{r}
#| label: 002-PP_logistic_fit-max_growth_rates.facet

PPfit + PPtime + plot_annotation(
  title = "***P. putida***: Time until max. growth rate for serial dilution of strating density", 
  tag_levels = "A") &
  theme(
    plot.title = element_markdown()
  )
```

```{r}
#| label: 002-BS_logistic_fit-max_growth_rates.facet

BSfit + BStime + plot_annotation(
  title = "***B. subtilis*: Time until max. growth rate for serial dilution of starting density**", 
  tag_levels = "A") &
  theme(
    plot.title = element_markdown()
  )
```

```{r}
#| label: 002-PP-BS-max_growth_rates.facet

PPtime + BStime + plot_annotation(
  title = "**Time until max. growth rate for serial dilution of starting density**", 
  tag_levels = "A") &
  theme(
    plot.title = element_markdown()
  )
```

