---
title: "Lotka-Volterra modeling three species interaction using the `gauseR` package"
author: "Leander Kühne, BSc"
date: "09/27/2026"
format: pdf
toc: true
toc-depth: 2
number-sections: true
execute:
  echo: false
  output: false
  warning: false
  cache: true
bibliography: references.bib
---

# Abstract


# Introduction

- Introduction (mainly include new facts that you needed to learn for this project, mainly exclude general background)

## Aim and project overview

This is a project for the "Practical Course in Modelling and Systems Biology" (2026S 300332-1). The aim of the project is to implement a Lotka-Volterra model using the `gauseR` to model a three species interaction experiment in spatial microbiology.

The experiment consists of a co-culturing experiment on microfluidic chips comprised with micro-scale mazes of different fractal complexity. The species used for co-culturing the two bacterial species *Pseudomonas putida* and *Bacillus subtilis* and the fungal species *Trichoderma ssp (?)*. All species are commonly found and co-isolated from soil habitats and are therefore thought to be coexisting in those ecosystems.

> I originally wanted to conduct the modeling analyses with my own data from co-culturing experiments with *P. putida*, *B. subtilis* and a complex soil bacterial community isolated from soil samples as a natural inoculum. However, since I am not having the necessary experimental data ready in time for the due date of this report I herein use the data of my colleague who conducted similar experiments with *Trichoderma* as the third species.

The overal aim is to see whether the experimental data can be properly represented by a Lotka-Voltera model and investigate if that approach is a viable method to narrow in on mechanistic insights to the interactions of this three species model in a spatially structured habitat.\
From the literature it is known, that P.p. and B.s. are frequently co-isolated from soils but cannot coexist in a well mixed environment under lab conditions. Prior experiments (not published) on microfluidic chips have shown that the introduction of a spatial structure allows for stable coexistence of these two species by niche differentiation.\
The underlying question of these experiments specifically are if *Trichoderma* alters the interactions between P.p and B.s. facilitating or hindering coexistence, by suppression or support of one or either of the species.\
On the broader scope of theoretical ecology the objective is to understand how community composition and (spatial) habitat complexity possibly facilitates coexistence allowing for the high diversity we can observe in nature. Micro-scale spatial structure is a key part of this framework since it is reducing interactions between (microbial) species (and organisms) and can thereby theoretically infinitely create niche differentiation overcoming potential fitness differences that would otherwise prohibit coexistence [@orrCoexistenceTheoryMicrobial2025]

In this setup we are tackling two shortcomings of most experimental microbial ecology research: Most research is focused on two species interactions exclusively but it is unclear how these results translate into more complex communities, since higher order interactions might be leading to emergent properties of microbial communities. Secondly, most experiments are carried out in well-mixed liquid cultures, omitting the importance of spatial structure as outlined above.\
Underrepresentation of these approaches might be one explanation to the unculturability problem that microbiological research faces since it's very beginings.

With the approach to reflect the interaction dynamics of this three species system in a simple Lotka-Volterra model I am trying to approximate on the dynamics with a simple model approach to see whether experimental data can be fitted satisfyingly on this degree of complexity already or if more sophisticated models need to be employed. Also the attempt to model the data with a LV model can give insights if the interactions can be described by transient competition or if higher-order interactions are at work.\
The Lotka-Volterra equation is while being a mathematically relatively straight forward model still a very powerful approach in ecology that can account for very complex and multi-layered network of interactions without the need to have a complete mechanistic understanding of these interactions on a molecular level.

This is the basic equation we are going to assume as Lotka-Volterra equation throughout is derived from @muhlbauerGauseRSimpleMethods2020:

Derived from @muhlbauerGauseRSimpleMethods2020 the basic Lotka-Volterra equation we are employing throughout this study describes the changes per unit time in the abundance of species $N_{i}$ as

$$
\frac{\mathrm{d}N_{i}}{\mathrm{d}t} = N_{i} (r_{i} + \mathrm{a}_{ii} N_{i} + \sum_{j} (\mathrm{a}_{ij} N_{j}))
$$ {#eq-lv}

where $r_{i}$ is the intrinsic growh rate... This is a reference to @fig-gfp-f5 which would be inline or in parentheses (@fig-gfp-f5). Moreover I already explained @eq-lv before.


Generalized LV:

$$
\frac{\mathrm{d}x_{i}(\mathbf{t})}{\mathrm{d}\mathbf{t}}=\mathbf{x}_{i}(\mathbf{t})\left(\mathbf{b}_{i}+\sum_{j=1}^{n}\mathbf{a}_{ij}x_{j}(\mathbf{t})\right)\qquad{i}=1,\ldots,n
$$



# Methods and Data
- Methods and data (provides reproducibility: 

## Experimental methods

Very brief and general.

## Software and Code

- data incl accession numbers+citation, 
- software incl version+non-default parameters+citation, 
- databases incl release number or release date + citation, 

## Data analysis and modeling workflow

- your workflows; BUT: usually no command lines


# Results
```{r}
# Load dependencies

library(tidyverse)
library(gauseR)
library(ggh4x)        # per-column coloured strips
library(kableExtra)   # PDF tables (booktabs, multi-level headers)
library(readxl)

options(knitr.kable.NA = "\u2013")   # missing values shown as "–"
# `summary_df` is assumed to be loaded already
# (columns: ch, time, fractal, species [= condition], mean_value, sd, n, se)
```

```{r}
# Import raw data and process

path <- "~/Documents/MSc_MMEI/MA/data/maryam_data"

df <- list.files(path, pattern = "^rep\\d+\\.xlsx$", full.names = TRUE) |>
  lapply(function(file) {
    data <- read_excel(file)
    data$rep <- as.integer(sub("rep(\\d+)\\.xlsx", "\\1", basename(file)))
    data
  }) |>
  bind_rows()

df <- setNames(df, c("position", "area", "mean", "min", "max", "ch", "time", "chip", "fractal", "rep"))

# Convert timepoints to days after inoculation
df$time <- df$time*2 

# Set Conditions
df$species <- dplyr::recode(
  df$chip,
  "1" = "T+B",
  "2" = "T+P",
  "3" = "T+P+B",
  "4" = "B",
  "5" = "P",
  "6" = "P+B",
  "7" = "T+B",
  "8" = "T+P",
  "9" = "T+P+B",
  "10" = "B",
  "11" = "P",
  "12" = "P+B",
  "13" = "P+B",
  "14" = "B",
  "15" = "T+B",
  "16" = "P",
  "17" = "T+P",
  "18" = "P+B",
  "19" = "T+P+B",
  "20" = "B",
  "21" = "T+B",
  "22" = "P",
  "23" = "T+P",
  "24" = "P+B",
  "25" = "T+P+B"
)

# Import CFU dataset
# CFU <- read_excel("~/Documents/MSc_MMEI/MA/data/maryam_data/CFU.xlsx")

# Calculate summary statistics.
summary_df <- df %>% 
  group_by(ch, time, fractal, species) %>%
  summarise( mean_value = mean(mean, na.rm = TRUE), 
             sd = sd(mean, na.rm = TRUE), n = n(), se = sd / sqrt(n), .groups = "drop" 
             )
```

```{r}
#| eval: false

# Export to csv. Future analyses can be carried out starting by importing the csv from below.
write.csv(df, "df_chips_maryam.csv", row.names = FALSE)
write.csv(summary_df, "df_chips_maryam.stat.csv", row.names = FALSE)

# Import CSV
df <- read.csv ("df_chips_maryam.csv")
summary_df <- read.csv ("df_chips_maryam.stat.csv")
```

```{r}
#| label: config

# Experimental Setup Configuration
#---------------------------------
#
# One row per species code that can occur in a condition string ("T+P+B").
#  - channel = NA  -> species is part of the community but NOT measured (T)
#  - name          -> used in legend and table headers
species_info <- tribble(
  ~species, ~channel, ~name,          ~colour,   ~linetype, ~shape,
  "P",      "G",      "P. putida",    "#1B7837", "solid",   16,
  "B",      "R",      "B. subtilis",  "#762A83", "42",      17,
  "T",      NA,       "T (name?)",    NA,        NA,        NA
)

# Which conditions appear as figure columns (in this order).
# Everything is fitted and goes into the tables regardless.
plot_conditions <- c("P", "B", "P+B", "T+P+B")

# Preferred order of conditions in the tables; conditions not listed are
# appended alphabetically, listed-but-absent ones are ignored.
condition_order <- c("P", "B", "P+B", "T+P", "T+B", "T+P+B")

# Column tints (recycled if you plot more conditions)
bg_palette <- c("#9CC3E6", "#E8A99A", "#9FD6AE", "#EBCB7F", "#C9B5E0", "#9ED3D3")

# ---- derived helpers (no edits needed) -------------------------------------
tr             <- species_info |> filter(!is.na(channel))
tracked        <- tr$species                       # fixed a_ij / r_i index order
species_colors <- set_names(tr$colour,   tr$species)
species_ltys   <- set_names(tr$linetype, tr$species)
species_shapes <- set_names(tr$shape,    tr$species)
species_names  <- set_names(tr$name,     tr$species)

split_cond <- function(x) strsplit(x, "+", fixed = TRUE)[[1]]
cond_short <- function(x) str_remove_all(x, "\\+")                 # "T+P+B" -> "TPB"
cond_full  <- function(x, fmt = c("plain", "latex")) {             # "P+B" -> "P. putida + B. subtilis"
  fmt <- match.arg(fmt)
  map_chr(x, \(cnd) {
    codes <- split_cond(cnd)
    nm <- coalesce(species_info$name[match(codes, species_info$species)], codes)
    if (fmt == "latex") nm <- paste0("\\textit{", nm, "}")
    paste(nm, collapse = " + ")
  })
}
```

```{r}
#| label: canonical

# 1. Canonical long layout
#-------------------------
#
# Every downstream step relies on this layout:
# `condition | fractal | time | species | mean_value | sd | se | n`
# (one row per condition x fractal x time x *measured* species).

canonical <- summary_df |>
  rename(condition = species) |>                                  # your "species" column holds the condition
  inner_join(tr |> select(species, channel), by = c("ch" = "channel")) |>  # channel -> species
  # keep only the channel(s) that belong to the species present in that condition
  filter(map2_lgl(species, condition, \(s, cnd) s %in% split_cond(cnd))) |>
  mutate(fractal = factor(paste0("f", fractal),
                          levels = paste0("f", sort(unique(fractal))))) |>
  select(condition, fractal, time, species, mean_value, sd, se, n) |>
  arrange(condition, fractal, species, time)

stopifnot(!anyDuplicated(canonical[c("condition", "fractal", "time", "species")]))

all_conditions <- intersect(union(condition_order, sort(unique(canonical$condition))),
                            unique(canonical$condition))
```

```{r}
#| label: fit
#| message: false

## 2. Fit every condition x fractal
#----------------------------------
#
# Fit one condition x fractal. Species columns are ordered like `tracked`,
# so the fitted index -> global index mapping is stable.
fit_gause <- function(d) {
  wide <- d |>
    select(time, species, mean_value) |>
    pivot_wider(names_from = species, values_from = mean_value) |>
    arrange(time)
  sp <- tracked[tracked %in% names(wide)]
  gause_wrapper(time = wide$time, species = as.data.frame(wide[sp]), doplot = FALSE)
}

fits <- canonical |>
  group_by(condition, fractal) |>
  nest() |>
  ungroup() |>
  mutate(
    sp  = map(data, \(d) tracked[tracked %in% d$species]),
    res = map(data, safely(fit_gause)),
    fit = map(res, "result"),
    err = map_chr(res, \(r) if (is.null(r$error)) NA_character_ else conditionMessage(r$error))
  )

failed <- filter(fits, !is.na(err))
if (nrow(failed)) {
  warning("Fits failed for: ",
          paste(failed$condition, failed$fractal, sep = "/", collapse = ", "), "\n",
          paste(unique(failed$err), collapse = "\n"))
}
fits_ok <- filter(fits, is.na(err))

# Diagnostic if something looks off (run interactively):
# str(fits_ok$fit[[1]], max.level = 1); fits_ok$fit[[1]]$parameter_intervals
```

```{r}
#| label: extract

# 3. Extract predictions, parameters and goodness of fit
#--------------------------------------------------------
#
extract_predicted <- function(fit, sp) {
  out <- as.data.frame(fit$out)
  out <- out[c("time", setdiff(names(out), "time"))]
  stopifnot(ncol(out) == length(sp) + 1)
  names(out) <- c("time", sp)                      # species order as passed to the fit
  pivot_longer(out, -time, names_to = "species", values_to = "N")
}

# ALL parameters (r_i and a_ij), re-indexed from "position within this fit"
# to the GLOBAL species index, so a B monoculture reports a22 / r2 (not a11 / r1).
extract_params <- function(fit, sp) {
  idx <- match(sp, tracked)
  # local index (position in this fit) -> global index, one name at a time
  # (scalar logic: no vector recycling that could go wrong)
  reindex <- function(p) {
    if (str_detect(p, "^r[0-9]$") || str_detect(p, "^a[0-9]{2}$")) {
      d <- as.integer(str_extract_all(p, "[0-9]")[[1]])
      paste0(str_sub(p, 1, 1), paste(idx[d], collapse = ""))
    } else p                                             # unknown names pass through unchanged
  }
  fit$parameter_intervals |>
    as.data.frame() |>
    rownames_to_column("parameter") |>
    mutate(parameter = map_chr(parameter, reindex),
           i = suppressWarnings(as.integer(str_sub(parameter, 2, 2))),   # NA for non r/a names
           j = suppressWarnings(as.integer(str_sub(parameter, 3, 3)))) |>  # NA for r_i
    select(parameter, i, j, lower_sd, mu, upper_sd)
}

predicted <- fits_ok |>
  mutate(pred = map2(fit, sp, extract_predicted)) |>
  select(condition, fractal, pred) |>
  unnest(pred)

params <- fits_ok |>
  mutate(par = map2(fit, sp, extract_params)) |>
  select(condition, fractal, par) |>
  unnest(par)

coefs <- filter(params, str_starts(parameter, "a"))   # interaction coefficients only (table 1)

observed <- canonical   # already in the right shape (incl. se for error bars)

# Goodness of fit per condition x fractal x species: predicted curve is
# interpolated at the observed time points and compared with the observed means.
gof <- observed |>
  semi_join(fits_ok, by = c("condition", "fractal")) |>
  group_by(condition, fractal, species) |>
  group_modify(\(d, key) {
    pr   <- filter(predicted, condition == key$condition,
                   fractal == key$fractal, species == key$species)
    yhat <- approx(pr$time, pr$N, xout = d$time)$y
    ok   <- is.finite(yhat)
    res  <- d$mean_value[ok] - yhat[ok]
    tibble(n_obs = sum(ok),
           rmse  = sqrt(mean(res^2)),
           r2    = 1 - sum(res^2) / sum((d$mean_value[ok] - mean(d$mean_value[ok]))^2))
  }) |>
  ungroup()
```

Here begins the text describing the results

First figure:

```{r}
#| output: true
#| label: fig-rfp-f5
#| fig-cap: RFP (solid line) and GFP (dashed line) flourescence intensity on microfluidic mazes with fractal order 5.


summary_df %>%
  ggplot(aes(x = time, color = species)) + 
  #geom_line(data=subset(summary_df, ch =='R'), aes(y = mean_value)) +
  geom_line(data=subset(summary_df, ch == 'G'& fractal == '5'), aes(y = mean_value), linetype = "dashed") +  
  geom_line(data=subset(summary_df, ch == 'R'& fractal == '5'), aes(y = mean_value)) +
  geom_errorbar(data=subset(summary_df, ch == 'G'& fractal == '5'), aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.3 ) + 
  geom_errorbar(data=subset(summary_df, ch == 'R'& fractal == '5'), aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.3 ) + 
  theme_light() +
  labs(y = "RFP/GFP intensity",
       x ="time [days]")
```


Second figure:

```{r}
#| label: fig-growth-grid
#| output: true
#| fig-width: 8
#| fig-height: 7
#| fig-cap: "Observed (points, ±SE) and predicted (lines) growth curves across fractal levels (rows) and conditions (columns)."

col_levels <- cond_short(plot_conditions)
bg <- set_names(rep_len(bg_palette, length(col_levels)), col_levels)

prep_plot <- function(df) {
  df |>
    filter(condition %in% plot_conditions) |>
    mutate(condition_label = factor(cond_short(condition), levels = col_levels))
}
predicted_p <- prep_plot(predicted)
observed_p  <- prep_plot(observed)

p <- ggplot() +
  # column background; no `fractal` column -> recycled into every row
  geom_rect(
    data = tibble(condition_label = factor(col_levels, levels = col_levels)),
    aes(xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, fill = condition_label),
    alpha = 0.5, show.legend = FALSE
  ) +
  geom_line(data = predicted_p,
            aes(time, N, colour = species, linetype = species), linewidth = 0.7) +
  geom_linerange(data = observed_p,
                 aes(time, ymin = mean_value - se, ymax = mean_value + se, colour = species),
                 linewidth = 0.3) +
  geom_point(data = observed_p,
             aes(time, mean_value, colour = species, shape = species), size = 1.6) +
  ggh4x::facet_grid2(
    fractal ~ condition_label,
    strip = ggh4x::strip_themed(
      background_x = ggh4x::elem_list_rect(fill = scales::alpha(unname(bg), 0.5)),
      background_y = ggh4x::elem_list_rect(fill = "grey20"),
      text_x = ggh4x::elem_list_text(face = "bold", colour = "grey15"),
      text_y = ggh4x::elem_list_text(face = "bold", colour = "white")
    )
  ) +
  scale_fill_manual(values = bg, guide = "none") +
  scale_colour_manual(values = species_colors,  labels = species_names, limits = tracked, name = "Species") +
  scale_linetype_manual(values = species_ltys,  labels = species_names, limits = tracked, name = "Species") +
  scale_shape_manual(values = species_shapes,   labels = species_names, limits = tracked, name = "Species") +
  scale_x_continuous(n.breaks = 3) +
  labs(x = "Time [days]", y = "Population size N [fluorescence intensity]") +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(colour = "grey88", linewidth = 0.25),
    panel.spacing.x  = unit(10, "pt"),
    panel.spacing.y  = unit(3,  "pt"),
    strip.placement  = "outside",
    axis.text        = element_text(size = 7, colour = "grey30"),
    axis.title       = element_text(size = 9.5),
    axis.ticks       = element_line(colour = "grey70", linewidth = 0.25),
    legend.position  = "bottom",
    legend.title     = element_text(size = 9, face = "bold"),
    legend.text      = element_text(size = 8),
    plot.margin      = margin(6, 6, 6, 6)
  )

p

# ggsave("growth_curve_figure.pdf", p, width = 6.5, height = 7, units = "in")
```


The table follows here. 

```{r}
#| label: tbl-lv-coeff
#| output: true
#| tbl-cap: "Estimated Lotka-Volterra interaction coefficients for cocultures of *P. putida* (1) and *B. subtilis* (2) on microfluidic chips with mazes of increasing fractal complexity (f0-f5). $a_{11}$ and $a_{22}$ are the intra-specific effects of *P. putida* and *B. subtilis*, respectively; $a_{12}$ and $a_{21}$ are the inter-specific effects of *B. subtilis* on *P. putida* and vice versa. Negative coefficients indicate competition/inhibition, positive coefficients promotion/enhancement."

tab_long <- coefs |>
  mutate(condition = factor(condition, levels = all_conditions),
         col = paste(condition, parameter, sep = "|")) |>
  arrange(condition, i != j, i, j)          # a11, a22, a12, a21 within each condition

tab_wide <- tab_long |>
  select(fractal, col, mu) |>
  pivot_wider(names_from = col, values_from = mu) |>   # column order = order of appearance
  arrange(fractal)

col_meta <- tab_long |>
  distinct(col, condition, parameter)
col_meta <- col_meta[match(names(tab_wide)[-1], col_meta$col), ]   # same order as table columns

grp <- count(col_meta, condition)                                  # columns per condition
top <- c(" " = 1, set_names(grp$n, cond_full(as.character(grp$condition), "latex")))

tab_wide |>
  kbl(format = "latex", booktabs = TRUE, digits = 4, align = "c", escape = FALSE,
      col.names = c("Fractal", sprintf("$a_{%s}$", str_sub(col_meta$parameter, 2)))) |>
  add_header_above(top, line = TRUE, escape = FALSE) |>   # rule under each condition group
  kable_styling(latex_options = "scale_down")
```


## Heatmap of interaction coefficients

```{r}
#| label: fig-heatmap
#| output: true
#| fig-width: 6.5
#| fig-height: 3.6
#| fig-cap: "Interaction coefficients per fractal level (rows) and condition (columns). Tile colour shows the estimated coefficient (blue: positive, red: negative); colour saturation shows $R^2$ of the fit of the species whose growth equation contains the coefficient (faint = poor fit). Numbers are the coefficient values."

# Conditions shown (default: all, like the table; e.g. use plot_conditions to restrict)
hm_conditions <- all_conditions

# column order within a condition, same as the table: a11, a22, a12, a21
coef_levels <- coefs |>
  distinct(parameter, i, j) |>
  arrange(i != j, i, j) |>
  pull(parameter)

heat <- coefs |>
  mutate(species = tracked[i]) |>              # a_ij belongs to the growth equation of species i
  left_join(select(gof, condition, fractal, species, r2),
            by = c("condition", "fractal", "species")) |>
  filter(condition %in% hm_conditions) |>
  mutate(condition_label = factor(cond_short(condition), levels = cond_short(hm_conditions)),
         parameter = factor(parameter, levels = coef_levels),
         r2_clamped = pmin(pmax(r2, 0), 1),      # negative R2 -> 0 (faintest)
         label = formatC(mu, format = "g", digits = 2))

m <- max(abs(heat$mu), na.rm = TRUE)             # symmetric limits: white = 0

sigma <- median(abs(heat$mu), na.rm = TRUE)

ggplot(heat, aes(x = parameter, y = fct_rev(fractal))) +
  geom_tile(aes(fill = mu, alpha = r2_clamped), colour = "white", linewidth = 0.6) +
  geom_text(aes(label = label), size = 2.2, colour = "grey15") +
  facet_grid(. ~ condition_label, scales = "free_x", space = "free_x") +
  scale_fill_gradient2(
    low = "#B2182B", mid = "white", high = "#2166AC",
    midpoint = 0, limits = c(-m, m), name = "Coefficient",
    transform = scales::pseudo_log_trans(sigma = sigma)) +
  scale_alpha_continuous(range = c(0.3, 1), limits = c(0, 1), name = expression(R^2)) +
  scale_x_discrete(position = "top",
                   labels = \(x) parse(text = sprintf("italic(a)[%s]", str_sub(x, 2)))) +
  guides(alpha = guide_legend(override.aes = list(fill = "grey30"))) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid      = element_blank(),
    strip.placement = "outside",                   # condition strip above the coefficient labels
    strip.background = element_rect(fill = "grey92", colour = NA),
    strip.text      = element_text(face = "bold", colour = "grey15"),
    panel.spacing.x = unit(8, "pt"),
    axis.text       = element_text(colour = "grey25"),
    legend.position = "bottom",
    legend.title    = element_text(size = 8, face = "bold"),
    legend.text     = element_text(size = 7)
  )
```

## Lotka—Volterra modeling using the `gauseR` library

### Parameterization and regression fit for *Pseudomonas* monoculture

```{r}
#| eval: false
##| output: true
# Subset of df with Pseudomonas monoculture for fractal complexity = 5
df_P <- subset(summary_df, species=="P" & ch=="G" & fractal=="5")

# calculate time-lagged abundance
lagged_data<-get_lag(x=df_P$mean_value, time = df_P$time)

# calculate per-capita growth rate
lagged_data$dNNdt <- percap_growth(x = lagged_data$x, laggedx = lagged_data$laggedx, dt = lagged_data$dt)

# plot relationship
plot(dNNdt~laggedx, lagged_data,
     xlab="Lagged Abundance (N)",
     ylab="Per-capita Growth Rate (dN/Ndt)")
abline(h=0, v=0, lty=3)

# fit model to relationship
mod<-lm(dNNdt~laggedx, lagged_data)
abline(mod, lwd=2, col=2)
```

Time-lagged data was generated using the `get_lag` function and the obtained data was used to calculate the per-capita growth rate.\
The per-capita growth rate for the logistic growth model was defined as:

$$ \frac{\mathrm{d}N}{\mathrm{d}t} / N = r - a_{ii} N $$

with $r$ being the intrinsic growth rate, and $a_{ii}$ the self-limitation in growth of the species.

The per-capita growth rate was plotted against the abundance and a linear regression was fitted to obtain the needed parameters $r$ (read as the y-intercept) and $K$ (read as the x-intercept). $a_{ii}$ is the slope of the linear regression.



```{r}
#| eval: false

# Eventually, $r$ and $K$ were used along with $N_{0}$ (abundance (fluorescence intensity) at time 0) to fit a logistic regression model using the `get logistic` function [@fig-P-mono-f5-logistic]. Coefficients were for the *P. putida* monoculture at maze fractal complexity of 5 are: $r =$ `{r} sprintf("%.3f", coef(mod)[1])`, $K =$ `{r} sprintf("%.3f",-coef(mod)[1]/coef(mod)[2])` and $N_{0} =$ `{r} sprintf("%.3f",print(df_P$mean_value)[1])`.


# r: y-intercept
r <- coef(mod)[1]

# K: x-intercept computed as 
K <- -coef(mod)[1]/coef(mod)[2]

# N0: observed abundance value (GFP intensity) at time 0
N0 <- print(df_P$mean_value)[1]

# Test goodness of fit
prediction_short<-get_logistic(time = df_P$time, N0 = (print(df_P$mean_value)[1]), r = (coef(mod)[1]), K= (-coef(mod)[1]/coef(mod)[2]))

goodness <- test_goodness_of_fit(observed = df_P$mean_value, predicted = prediction_short)
```

```{r}
#| eval: false
##| output: true
#| label: fig-P-mono-f5-logistic
#| fig-cap: Logistic growth curve fitted from parameterization of per-capita growth rate. The plot shown here is exemplary and data is from *P. putida* monoculture at maze fractal complexity of 5. Parameters to fit the linear regression are $r$ is , $K =$  and $N_{0}$ is the observed abundance value (GFP intensity) at time 0:.

# Logistic fit
plot(mean_value~time, df_P, xlab="Day", ylab="GFP intensity of P. putida monoculture")

timelst<-seq(0, 22, by=0.1) # sequence of time values to plot
prediction<-get_logistic(time = timelst, N0 = (print(df_P$mean_value)[1]), r = (coef(mod)[1]), K= (-coef(mod)[1]/coef(mod)[2]))

lines(timelst, prediction, lwd=2)
```

```{r}
#| eval: false
#| message: false
# Extraction of interaction coefficients

wide_df <- summary_df %>%
  pivot_wider(names_from = ch, values_from = c(mean_value, sd, se), names_prefix = "N_")

extract_interactions <- function(result, fractal) {
  
  result$parameter_intervals %>%
    as.data.frame() %>%
    tibble::rownames_to_column("parameter") %>%
    filter(parameter %in% c("a11", "a12", "a21", "a22")) %>%
    mutate(
      fractal = fractal
    ) %>%
    select(fractal, parameter, lower_sd, mu, upper_sd)
}


interaction_results <- lapply(
  names(gause_results),
  \(f) extract_interactions(
    gause_results[[f]],
    f
  )
) %>%
  bind_rows()

interaction_table <- interaction_results %>%
  select(fractal, parameter, mu) %>%
  pivot_wider(
    names_from = parameter,
    values_from = mu
  )

```

```{r}
#| eval: false
##| output: true
#| label: tbl-coef-BP
#| tbl-cap: Estimated Lotka-Volterra interaction coefficients for a coculture of *P. putida* (1) and *B. subtilis* (2) on microfluidic chips with mazes of incresing fractal complexity (f0-5). $a_{11}$, $a_{22}$ describe the intra-specific effect of *P. putida* and *B. subtilis* respectively. $a_{12}$ and $a_{21} describe the inter-specific effect of *B. subtilis* on *P. putida* and the other way around, respectively. Negative coefficients resemble competition/inhibition whereas positive coefficients resemble promotion/enhancement.

# Interaction coefficients table
knitr::kable(
  interaction_table,
  digits = 4,
  col.names = c(
    "Fractal",
    "$a_{11}$",
    "$a_{12}$",
    "$a_{21}$",
    "$a_{22}$"
  ))
```

# Discussion

- Discussion (did we succeed w.r.t. the project aims? did we make particular assumptions? do our methods have weaknesses? how do the results compare to previous expectations and previous work)


# Statement of authorship and use of AI and LLMs

This report is submitted as the practical project for the class "Practial Course in Modelling and Systems Biology" (2026S 300332-1). The used experimental data has been gathered and provided by Miryam Al-Jalahwee, BSc. All data arranging, sorting and subsequent analysis has been conducted by the author. 


# References

::: {#refs}
:::

# Supplementary Material

- Supplementary material: code and command lines (if needed)
```{r}
#| label: tbl-gof
#| output: true
#| tbl-cap: "Goodness of fit of the fitted growth curves against the observed means, per species. $R^2 = 1 - SS_\\mathrm{res}/SS_\\mathrm{tot}$ (can be negative for fits worse than the mean); RMSE in the units of $N$ (fluorescence intensity). Model predictions were interpolated at the observed time points."

sp_present <- tracked[tracked %in% gof$species]
gof_cols   <- as.vector(t(outer(sp_present, c("r2", "rmse"), paste, sep = "|")))

gof_wide <- gof |>
  mutate(condition = factor(condition, levels = all_conditions),
         species   = factor(species, levels = sp_present)) |>
  arrange(condition, fractal) |>
  pivot_wider(id_cols = c(condition, fractal), names_from = species,
              values_from = c(r2, rmse), names_glue = "{species}|{.value}") |>
  select(condition, fractal, all_of(gof_cols)) |>
  mutate(condition = cond_full(as.character(condition), "latex"),
         fractal   = as.character(fractal))

gof_wide |>
  kbl(format = "latex", booktabs = TRUE, digits = 3, align = "llcccc", escape = FALSE,
      col.names = c("Condition", "Fractal", rep(c("$R^2$", "RMSE"), length(sp_present)))) |>
  add_header_above(c(" " = 2,
                     set_names(rep(2, length(sp_present)),
                               paste0("\\textit{", species_names[sp_present], "}"))),
                   line = TRUE, escape = FALSE) |>
  collapse_rows(columns = 1, latex_hline = "major", valign = "top") |>
  kable_styling(font_size = 9)
```

```{r}
#| label: tbl-confidence
#| output: true
#| tbl-cap: "Parameter estimates with uncertainty for every fit. $r_i$: intrinsic growth rate of species $i$; $a_{ij}$: interaction coefficients (indices as in Table 1). SD is half the width of the ±1 SD interval returned by gauseR; the 95\\% interval is the normal approximation (estimate ± 1.96 SD). 'Excludes 0' indicates whether this interval excludes zero."

param_out <- params |>
  mutate(
    condition = factor(condition, levels = all_conditions),
    sd = abs(upper_sd - lower_sd) / 2,
    lo = mu - 1.96 * sd,
    hi = mu + 1.96 * sd
  ) |>
  arrange(condition, fractal, str_starts(parameter, "a"), coalesce(i != j, FALSE), parameter) |>
  transmute(
    Condition = cond_full(as.character(condition), "latex"),
    Fractal   = as.character(fractal),
    Parameter = sprintf("$%s_{%s}$", str_sub(parameter, 1, 1), str_sub(parameter, 2)),
    Estimate  = sprintf("%.4f", mu),
    SD        = sprintf("%.4f", sd),
    Interval  = sprintf("[%.4f, %.4f]", lo, hi),
    Excl0     = if_else(lo > 0 | hi < 0, "yes", "no")
  )

param_out |>
  kbl(format = "latex", booktabs = TRUE, longtable = TRUE, escape = FALSE,
      align = "llcrrrc", row.names = FALSE,
      col.names = c("Condition", "Fractal", "Parameter", "Estimate", "SD",
                    "95\\% interval", "Excludes 0")) |>
  collapse_rows(columns = 1:2, latex_hline = "major", valign = "top") |>
  kable_styling(latex_options = "repeat_header", font_size = 8)
```
