---
title: "003-NI-dil_series-media_test"
format: html
identifier: "003"
filters:
   - lightbox
lightbox: auto
---

# Short Description

This experiment tried to investigate in the effect of pre-culturing the soil samples in LB and M9 media, respectively on growth rates of subsequently isolated natural soil communities.

1 g of soil was incubated overnight in either LB or M9. 4 soil replicates for each medium. After o/n culture media were vacuum filtered to obtain bacterial communities and serial dilutions ranging from 10^-2^ to 10^-6^ were incubated on 96-well-plates (8 technical replicates). For \~1.5 days.

# Data loading and `df` arrangement

```{r}
#| output: false

library(tidyverse)
library(readxl)
library(data.table)
library(tibble)
library(patchwork)
library(ggnested)
library(ggtext)
```

# Plate 1

Repeat this for all 4 Plates.

```{r}
#| output: false

# Get files 
files <- list.files(
  path = "~/Documents/MSc_MMEI/MA/data/003-Soil_dil_series",
  pattern = "*.xlsx",
  full.names = TRUE,
)

data <- lapply(files, function(file_path) {
  readxl::read_excel(file_path, sheet = "Plate1")
})

## Plate Processing Pipeline (w/o sheets) 
process_plate <- function(file_path) {
  
  df <- read_excel(file_path, sheet = "Plate1", col_names = FALSE)
  
  # Load measurement values from excel
  OD600 <- df[25:32,2:13]
  GFP100 <- df[57:64,2:13]
  GFP80 <- df[89:96,2:13]
  RFP100 <- df[121:128,2:13]
  RFP80 <- df[153:160,2:13]
  
  # Convert to long format
  OD600 <- as.data.frame(as.table(as.matrix(OD600)))
  GFP100 <- as.data.frame(as.table(as.matrix(GFP100)))
  GFP80 <- as.data.frame(as.table(as.matrix(GFP80)))
  RFP100 <- as.data.frame(as.table(as.matrix(RFP100)))
  RFP80 <- as.data.frame(as.table(as.matrix(RFP80)))
  
  colnames(OD600) <- c("row", "col", "value") # Rename columns
  colnames(GFP100) <- c("row", "col", "value") # Rename columns
  colnames(GFP80) <- c("row", "col", "value") # Rename columns
  colnames(RFP100) <- c("row", "col", "value") # Rename columns
  colnames(RFP80) <- c("row", "col", "value") # Rename columns
  
  # Fix columns
  OD600$col <- as.numeric(OD600$col) 
  GFP100$col <- as.numeric(GFP100$col) 
  GFP80$col  <- as.numeric(GFP80$col) 
  RFP100$col <- as.numeric(RFP100$col) 
  RFP80$col  <- as.numeric(RFP80$col) 
  
  # Add wavelength
  OD600$wavelength <- "OD600"
  GFP100$wavelength <- "GFP100"
  GFP80$wavelength <- "GFP80"
  RFP100$wavelength <- "RFP100"
  RFP80$wavelength <- "RFP80"
  
  df_clean <- rbindlist(list(OD600, GFP100, GFP80, RFP100, RFP80)) # Rbind (concatenate) data.tables
  
  df_clean$value <- as.numeric(as.character(df_clean$value)) # Convert values to numeric
  
  df_clean$well <- paste0(df_clean$row, df_clean$col) # Add well ID
  
  df_clean$media <- "LB"
  
  # Add soil sample
  df_clean$sample <- dplyr::recode(
    df_clean$col,
    "1" = "1",
    "2" = "1",
    "3" = "1",
    "4" = "1",
    "5" = "1",
    "6" = "1",
    "7" = "2",
    "8" = "2",
    "9" = "2",
    "10" = "2",
    "11" = "2",
    "12" = "2"
  )
    
  # Add dilution
  df_clean$dilution <- dplyr::recode(
    df_clean$col,
    "1" = "10^-3",
    "2" = "10^-4",
    "3" = "10^-5",
    "4" = "10^-6",
    "5" = "ctrl",
    "6" = "ctrl",
    "7" = "10^-3",
    "8" = "10^-4",
    "9" = "10^-5",
    "10" = "10^-6",
    "11" = "ctrl",
    "12" = "ctrl",
  )
  
  # Extract date and time from excel 
  # Extract metadata (date + time)
  date <- read_excel(file_path, sheet = "Plate1", range = "B5", col_names = FALSE)[[1]]
  time <- read_excel(file_path, sheet = "Plate1", range = "B6", col_names = FALSE)[[1]]
  
  # Combine into datetime
  datetime <- as.POSIXct(
    paste(date, time),
    format = "%Y-%m-%d %I:%M:%S %p"
  )
  
  # Add datetime and filename to dataframe
  df_clean <- df_clean %>%
    dplyr::mutate(
      datetime = datetime,
      file = basename(file_path)
    )
  
  return(df_clean)
}

## Execute function and combine dataset
combined_data<- lapply(files, process_plate) %>%
  data.table::rbindlist()

# Add elapsed time in hours (first value = 0)
combined_data$elapsed_hours <- as.numeric(
  difftime(combined_data$datetime, combined_data$datetime[1], units = "hours")
)

# Tidy dataframe
combined_data <- subset(combined_data, select = -c(row, col)) # Remove date columns
combined_data <- combined_data[, c("file", "well", "datetime", "elapsed_hours", "sample", "media", "dilution", "wavelength", "value")] # Reorder

P1_df <- combined_data
# Summary statistics
P1_summary_stat <- combined_data %>%
  group_by(sample, dilution, wavelength, elapsed_hours, media, datetime, file) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n),
    .groups = "drop"
  )
```

## Plate 2

```{r}
#| output: false
#| echo: false

data <- lapply(files, function(file_path) {
  readxl::read_excel(file_path, sheet = "Plate2")
})

## Plate Processing Pipeline (w/o sheets)
process_plate <- function(file_path) {
  
  df <- read_excel(file_path, sheet = "Plate2", col_names = FALSE)
  
  # Load measurement values from excel
  OD600 <- df[25:32,2:13]
  GFP100 <- df[57:64,2:13]
  GFP80 <- df[89:96,2:13]
  RFP100 <- df[121:128,2:13]
  RFP80 <- df[153:160,2:13]
  
  # Convert to long format
  OD600 <- as.data.frame(as.table(as.matrix(OD600)))
  GFP100 <- as.data.frame(as.table(as.matrix(GFP100)))
  GFP80 <- as.data.frame(as.table(as.matrix(GFP80)))
  RFP100 <- as.data.frame(as.table(as.matrix(RFP100)))
  RFP80 <- as.data.frame(as.table(as.matrix(RFP80)))
  
  colnames(OD600) <- c("row", "col", "value") # Rename columns
  colnames(GFP100) <- c("row", "col", "value") # Rename columns
  colnames(GFP80) <- c("row", "col", "value") # Rename columns
  colnames(RFP100) <- c("row", "col", "value") # Rename columns
  colnames(RFP80) <- c("row", "col", "value") # Rename columns
  
  # Fix columns
  OD600$col <- as.numeric(OD600$col) 
  GFP100$col <- as.numeric(GFP100$col) 
  GFP80$col  <- as.numeric(GFP80$col) 
  RFP100$col <- as.numeric(RFP100$col) 
  RFP80$col  <- as.numeric(RFP80$col) 
  
  # Add wavelength
  OD600$wavelength <- "OD600"
  GFP100$wavelength <- "GFP100"
  GFP80$wavelength <- "GFP80"
  RFP100$wavelength <- "RFP100"
  RFP80$wavelength <- "RFP80"
  
  df_clean <- rbindlist(list(OD600, GFP100, GFP80, RFP100, RFP80)) # Rbind (concatenate) data.tables
  
  df_clean$value <- as.numeric(as.character(df_clean$value)) # Convert values to numeric
  
  df_clean$well <- paste0(df_clean$row, df_clean$col) # Add well ID
  
  df_clean$media <- "LB"
  
  # Add soil sample
  df_clean$sample <- dplyr::recode(
    df_clean$col,
    "1" = "3",
    "2" = "3",
    "3" = "3",
    "4" = "3",
    "5" = "3",
    "6" = "3",
    "7" = "4",
    "8" = "4",
    "9" = "4",
    "10" = "4",
    "11" = "4",
    "12" = "4"
  )
  
  # Add dilution
  df_clean$dilution <- dplyr::recode(
    df_clean$col,
    "1" = "10^-2",
    "2" = "10^-3",
    "3" = "10^-4",
    "4" = "10^-5",
    "5" = "10^-6",
    "6" = "ctrl",
    "7" = "10^-2",
    "8" = "10^-3",
    "9" = "10^-4",
    "10" = "10^-5",
    "11" = "10^-6",
    "12" = "ctrl",
  )
  
  # Extract date and time from excel 
  # Extract metadata (date + time)
  date <- read_excel(file_path, sheet = "Plate2", range = "B5", col_names = FALSE)[[1]]
  time <- read_excel(file_path, sheet = "Plate2", range = "B6", col_names = FALSE)[[1]]
  
  # Combine into datetime
  datetime <- as.POSIXct(
    paste(date, time),
    format = "%Y-%m-%d %I:%M:%S %p"
  )
  
  # Add datetime and filename to dataframe
  df_clean <- df_clean %>%
    dplyr::mutate(
      datetime = datetime,
      file = basename(file_path)
    )

  return(df_clean)
}

## Execute function and combine dataset
combined_data<- lapply(files, process_plate) %>%
  data.table::rbindlist()

# Add elapsed time in hours (first value = 0)
combined_data$elapsed_hours <- as.numeric(
  difftime(combined_data$datetime, combined_data$datetime[1], units = "hours")
)

# Tidy dataframe
combined_data <- subset(combined_data, select = -c(row, col)) # Remove date columns
combined_data <- combined_data[, c("file", "well", "datetime", "elapsed_hours", "sample", "media", "dilution", "wavelength", "value")] # Reorder

P2_df <- combined_data

# Summary statistics
P2_summary_stat <- combined_data %>%
  group_by(sample, dilution, wavelength, elapsed_hours, media, datetime, file) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n),
    .groups = "drop"
  )
```

## Plate 3

```{r}
#| output: false
#| echo: false

# Get files
files <- list.files(
  path = "~/Documents/MSc_MMEI/MA/data/003-Soil_dil_series",
  pattern = "*.xlsx",
  full.names = TRUE,
)

data <- lapply(files, function(file_path) {
  readxl::read_excel(file_path, sheet = "Plate3")
})

## Plate Processing Pipeline (w/o sheets)
process_plate <- function(file_path) {
  
  df <- read_excel(file_path, sheet = "Plate3", col_names = FALSE)
  
  # Load measurement values from excel
  OD600 <- df[25:32,2:13]
  GFP100 <- df[57:64,2:13]
  GFP80 <- df[89:96,2:13]
  RFP100 <- df[121:128,2:13]
  RFP80 <- df[153:160,2:13]
  
  # Convert to long format
  OD600 <- as.data.frame(as.table(as.matrix(OD600)))
  GFP100 <- as.data.frame(as.table(as.matrix(GFP100)))
  GFP80 <- as.data.frame(as.table(as.matrix(GFP80)))
  RFP100 <- as.data.frame(as.table(as.matrix(RFP100)))
  RFP80 <- as.data.frame(as.table(as.matrix(RFP80)))
  
  colnames(OD600) <- c("row", "col", "value") # Rename columns
  colnames(GFP100) <- c("row", "col", "value") # Rename columns
  colnames(GFP80) <- c("row", "col", "value") # Rename columns
  colnames(RFP100) <- c("row", "col", "value") # Rename columns
  colnames(RFP80) <- c("row", "col", "value") # Rename columns
  
  # Fix columns
  OD600$col <- as.numeric(OD600$col) 
  GFP100$col <- as.numeric(GFP100$col) 
  GFP80$col  <- as.numeric(GFP80$col) 
  RFP100$col <- as.numeric(RFP100$col) 
  RFP80$col  <- as.numeric(RFP80$col) 
  
  # Add wavelength
  OD600$wavelength <- "OD600"
  GFP100$wavelength <- "GFP100"
  GFP80$wavelength <- "GFP80"
  RFP100$wavelength <- "RFP100"
  RFP80$wavelength <- "RFP80"
  
  df_clean <- rbindlist(list(OD600, GFP100, GFP80, RFP100, RFP80)) # Rbind (concatenate) data.tables
  
  df_clean$value <- as.numeric(as.character(df_clean$value)) # Convert values to numeric
  
  df_clean$well <- paste0(df_clean$row, df_clean$col) # Add well ID
  
  df_clean$media <- "M9"
  
  # Add soil sample
  df_clean$sample <- dplyr::recode(
    df_clean$col,
    "1" = "5",
    "2" = "5",
    "3" = "5",
    "4" = "5",
    "5" = "5",
    "6" = "5",
    "7" = "6",
    "8" = "6",
    "9" = "6",
    "10" = "6",
    "11" = "6",
    "12" = "6"
  )
  
  # Add dilution
  df_clean$dilution <- dplyr::recode(
    df_clean$col,
    "1" = "10^-2",
    "2" = "10^-3",
    "3" = "10^-4",
    "4" = "10^-5",
    "5" = "10^-6",
    "6" = "ctrl",
    "7" = "10^-2",
    "8" = "10^-3",
    "9" = "10^-4",
    "10" = "10^-5",
    "11" = "10^-6",
    "12" = "ctrl",
  )
  
  # Extract date and time from excel 
  # Extract metadata (date + time)
  date <- read_excel(file_path, sheet = "Plate3", range = "B5", col_names = FALSE)[[1]]
  time <- read_excel(file_path, sheet = "Plate3", range = "B6", col_names = FALSE)[[1]]
  
  # Combine into datetime
  datetime <- as.POSIXct(
    paste(date, time),
    format = "%Y-%m-%d %I:%M:%S %p"
  )
  
  # Add datetime and filename to dataframe
  df_clean <- df_clean %>%
    dplyr::mutate(
      datetime = datetime,
      file = basename(file_path)
    )

  return(df_clean)
}

## Execute function and combine dataset
combined_data<- lapply(files, process_plate) %>%
  data.table::rbindlist()

# Add elapsed time in hours (first value = 0)
combined_data$elapsed_hours <- as.numeric(
  difftime(combined_data$datetime, combined_data$datetime[1], units = "hours")
)

# Tidy dataframe
combined_data <- subset(combined_data, select = -c(row, col)) # Remove date columns
combined_data <- combined_data[, c("file", "well", "datetime", "elapsed_hours", "sample", "media", "dilution", "wavelength", "value")] # Reorder

P3_df <- combined_data

# Summary statistics
P3_summary_stat <- combined_data %>%
  group_by(sample, dilution, wavelength, elapsed_hours, media, datetime, file) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n),
    .groups = "drop"
  )
```

## Plate 4

```{r}
#| output: false
#| echo: false

# Get files
files <- list.files(
  path = "~/Documents/MSc_MMEI/MA/data/003-Soil_dil_series",
  pattern = "*.xlsx",
  full.names = TRUE,
)

data <- lapply(files, function(file_path) {
  readxl::read_excel(file_path, sheet = "Plate4")
})

## Plate Processing Pipeline (w/o sheets)
process_plate <- function(file_path) {
  
  df <- read_excel(file_path, sheet = "Plate4", col_names = FALSE)
  
  # Load measurement values from excel
  OD600 <- df[25:32,2:13]
  GFP100 <- df[57:64,2:13]
  GFP80 <- df[89:96,2:13]
  RFP100 <- df[121:128,2:13]
  RFP80 <- df[153:160,2:13]
  
  # Convert to long format
  OD600 <- as.data.frame(as.table(as.matrix(OD600)))
  GFP100 <- as.data.frame(as.table(as.matrix(GFP100)))
  GFP80 <- as.data.frame(as.table(as.matrix(GFP80)))
  RFP100 <- as.data.frame(as.table(as.matrix(RFP100)))
  RFP80 <- as.data.frame(as.table(as.matrix(RFP80)))
  
  colnames(OD600) <- c("row", "col", "value") # Rename columns
  colnames(GFP100) <- c("row", "col", "value") # Rename columns
  colnames(GFP80) <- c("row", "col", "value") # Rename columns
  colnames(RFP100) <- c("row", "col", "value") # Rename columns
  colnames(RFP80) <- c("row", "col", "value") # Rename columns
  
  # Fix columns
  OD600$col <- as.numeric(OD600$col) 
  GFP100$col <- as.numeric(GFP100$col) 
  GFP80$col  <- as.numeric(GFP80$col) 
  RFP100$col <- as.numeric(RFP100$col) 
  RFP80$col  <- as.numeric(RFP80$col) 
  
  # Add wavelength
  OD600$wavelength <- "OD600"
  GFP100$wavelength <- "GFP100"
  GFP80$wavelength <- "GFP80"
  RFP100$wavelength <- "RFP100"
  RFP80$wavelength <- "RFP80"
  
  df_clean <- rbindlist(list(OD600, GFP100, GFP80, RFP100, RFP80)) # Rbind (concatenate) data.tables
  
  df_clean$value <- as.numeric(as.character(df_clean$value)) # Convert values to numeric
  
  df_clean$well <- paste0(df_clean$row, df_clean$col) # Add well ID
  
  df_clean$media <- "M9"
  
  # Add soil sample
  df_clean$sample <- dplyr::recode(
    df_clean$col,
    "1" = "7",
    "2" = "7",
    "3" = "7",
    "4" = "7",
    "5" = "7",
    "6" = "7",
    "7" = "8",
    "8" = "8",
    "9" = "8",
    "10" = "8",
    "11" = "8",
    "12" = "8"
  )
  
  # Add dilution
  df_clean$dilution <- dplyr::recode(
    df_clean$col,
    "1" = "10^-2",
    "2" = "10^-3",
    "3" = "10^-4",
    "4" = "10^-5",
    "5" = "10^-6",
    "6" = "ctrl",
    "7" = "10^-2",
    "8" = "10^-3",
    "9" = "10^-4",
    "10" = "10^-5",
    "11" = "10^-6",
    "12" = "ctrl",
  )
  
  # Extract date and time from excel 
  # Extract metadata (date + time)
  date <- read_excel(file_path, sheet = "Plate4", range = "B5", col_names = FALSE)[[1]]
  time <- read_excel(file_path, sheet = "Plate4", range = "B6", col_names = FALSE)[[1]]
  
  # Combine into datetime
  datetime <- as.POSIXct(
    paste(date, time),
    format = "%Y-%m-%d %I:%M:%S %p"
  )
  
  # Add datetime and filename to dataframe
  df_clean <- df_clean %>%
    dplyr::mutate(
      datetime = datetime,
      file = basename(file_path)
    )

  return(df_clean)
}

## Execute function and combine dataset
combined_data<- lapply(files, process_plate) %>%
  data.table::rbindlist()

# Add elapsed time in hours (first value = 0)
combined_data$elapsed_hours <- as.numeric(
  difftime(combined_data$datetime, combined_data$datetime[1], units = "hours")
)

# Tidy dataframe
combined_data <- subset(combined_data, select = -c(row, col)) # Remove date columns
combined_data <- combined_data[, c("file", "well", "datetime", "elapsed_hours", "sample", "media", "dilution", "wavelength", "value")] # Reorder

P4_df <- combined_data

# Summary statistics
P4_summary_stat <- combined_data %>%
  group_by(sample, dilution, wavelength, elapsed_hours, media, datetime, file) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n),
    .groups = "drop"
  )
```

# Combine dataframes from different 96-well-plates

```{r}
df <- rbind(P1_df, P2_df, P3_df, P4_df)

df$sample <- as.character(df$sample) # Convert values to numeric

df <- df %>%
  mutate(elapsed_hours = round(elapsed_hours * 2) / 2)  # bins to 0.5 h
```

# Calculate summary statistics (mean and standard deviation)

```{r}
df_stat <- df %>%
  group_by(sample, dilution, wavelength, elapsed_hours, media, datetime, file) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n),
    .groups = "drop"
  )

# Mean and sd by replicates AND media replicates \
# (different soil samples) for LB/M9 media effect contribution
media_stat <- df %>%
  group_by(dilution, wavelength, elapsed_hours, media) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    n = n(),
    se = sd / sqrt(n),
    .groups = "drop"
  )
```

## Write/Load CSV

```{r}
write.csv(df, "df.csv", row.names = FALSE)
write.csv(df_stat, "df_stat.csv", row.names = FALSE)
write.csv(media_stat, "media_stat.csv", row.names = FALSE)
```

**Load `df`'s from CSV:**

```{r}
df <- read.csv ("df.csv") 
df_stat <- read.csv("df_stat.csv")
media_stat <- read.csv ("media_stat.csv") 
```

# Data Visualization

## OD600 10^-3^ all samples

```{r}
#| label: 003-NI-diff_media_OD

df_stat %>% 
  filter(wavelength %in% c("OD600") &
           dilution %in% c("10^-3")) %>%
  ggplot(aes(x = elapsed_hours, y = mean_value, color = sample)) +
  geom_line() +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_value - se, ymax = mean_value + se),
    width = 0.5 ) +
  # scale_color_discrete(labels = parse(text = unique(df_stat$dilution))) +
  theme_minimal() +
  labs(y= expression("OD"[600]), x="Time", title = expression("OD"[600]*" of all 10"^"-3"*" Dilutions"))

## OD600 10^-2 by media


```

```{r}
#| label: 003-OD_grouped_by_media.facet

media_stat %>% 
  filter(wavelength %in% c("OD600")) %>%
  ggplot(aes(x = elapsed_hours, y = mean_value, color = media)) +
  geom_line() +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_value - se, ymax = mean_value + se),
    width = 0.5 ) +
  # scale_color_discrete(labels = parse(text = unique(df_stat$dilution))) +
  theme_minimal() +
  labs(y= expression("OD"[600]), x="Time", title = expression("OD"[600]*" of all Dilutions grouped by media type")) +
  theme(plot.caption = element_markdown()) +
  facet_wrap(~dilution)
```

## OD600 all samples, facet dilutions

```{r}
#| label: 003-OD_by_different_dil.facet

df_stat %>% 
  filter(wavelength %in% c("OD600")) %>%
  ggplot(aes(x = elapsed_hours, y = mean_value, color = sample)) +
  geom_line() +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_value - se, ymax = mean_value + se),
    width = 0.5 ) +
  theme_minimal() +
  labs(y= expression("OD"[600]), x="Time", title = expression("OD"[600]*" Of All Soil Samples At Different Starting Dilutions")) +
  facet_wrap(~ dilution)
```

## Plotting for LB and M9

```{r}
#| label: 003-OD_grouped_by_media_and_dil.facet

media_stat %>% 
  filter(wavelength %in% c("OD600")) %>%
  ggplot(aes(x = elapsed_hours, y = mean_value, color = media)) +
  geom_line() +
  geom_point() +
  geom_errorbar(
    aes(ymin = mean_value - se, ymax = mean_value + se),
    width = 0.5 ) +
  theme_minimal() +
  labs(y= expression("OD"[600]), x="Time", title = expression("OD"[600]*" Of All Soil Samples At Different Starting Dilutions")) +
  facet_wrap(~ dilution)
```

## OD all grouped by media and dilution

```{r}
#| label: 003-OD_grouped_by_media_and_dil.ggnested

media_stat %>% filter(wavelength %in% c("OD600")) %>% 
  ggnested(
    aes(x = elapsed_hours, y = mean_value, main_group = media, sub_group = dilution), 
    legend_title = "Starting Density", 
    main_keys = T, 
    base_clr = "darkgreen", 
    gradient_type = "tints" ) +
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.2 ) + 
  theme_minimal() + 
  labs(y= expression("OD"[600]*" absorbance"), x="Time [h]", title = "OD<sub>600</sub> Growth Curves by Media Type and Serial Dilution of Starting Density") + 
  theme(plot.title = element_markdown())
```

## Everything else

```{r}
#| label: 003-OD_10^-3_by_media

df_stat %>% filter(dilution == "10^-3" & wavelength == "OD600") %>% 
  ggnested(
    aes(x = elapsed_hours, y = mean_value, main_group = media, sub_group = sample), 
    legend_title = "Starting Density", 
    main_keys = T, 
    base_clr = "darkgreen", 
    gradient_type = "tints" ) +
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.2 ) + 
  theme_minimal() + 
  labs(
    y= expression("OD"[600]), 
    x="Time [h]", 
    title = "OD<sub>600</sub> Growth Curves by Media Type for 10<sup>-3</sup> Starting Density") + 
  theme(plot.title = element_markdown())
```
