---
title: "008-Media_replenishment"
format: html
identifier: "004"
lightbox: true
---

# Short description
In this experiment 7 different conditions — consisting of all possible combinations of *P. putida*, *B. subtilis* and a natural soil inoculum — were (co-)cultured in a well mixed environment (96-well-plate). Every 48 h the cultures were diluted 1:5 into fresh M9 media and further used as the new main culture. 
12 biological replicates. 
After the third cycle, cross-contamination between conditions and replicates was introduced and experiment was subsequently aborted. 

# Data Loading and Process Function

```{r}
#| echo: false
#| output: false
#| label: 8527a891

library(tidyverse)
library(readxl)
library(stringr)
library(patchwork)
library(RColorBrewer)
library(ggnested)
library(growthcurver)
library(ggtext)
library(plotly)
library(data.table)
```


```{r}
#| label: e29c9c18
process_plate <- function(file_path) {
  
  raw <- read_excel(file_path, col_names = FALSE)
  
  # Extract plate sections
  OD600  <- raw[37:44, 2:13]
  GFP70 <- raw[70:77, 2:13]
  RFP85 <- raw[103:110, 2:13]
  RFP90  <- raw[136:143, 2:13]
  
  # Convert to long format
  to_long <- function(df, name) {
    out <- as.data.frame(as.table(as.matrix(df)))
    out$wavelength <- name
    return(out)
  }
  
  OD600  <- to_long(OD600,  "OD600")
  GFP70 <- to_long(GFP70, "GFP70")
  RFP85 <- to_long(RFP85, "RFP85")
  RFP90  <- to_long(RFP90,  "RFP90")
  
  # Combine
  plate_concat <- rbindlist(list(OD600, GFP70, RFP85, RFP90))
  setnames(plate_concat, c("row", "col", "value", "wavelength"))
  
  plate_concat$value <- as.numeric(as.character(plate_concat$value))
  plate_concat$well <- paste0(plate_concat$row, plate_concat$col)
  
  # Date + time from same file
  date <- read_excel(file_path, range = "E6", col_names = FALSE)[[1]]
  time <- read_excel(file_path, range = "E7", col_names = FALSE)[[1]]
  
  plate_concat$datetime <- as.POSIXct(paste(date, time))
  
  # Conditions
  plate_concat$condition <- dplyr::recode(
    plate_concat$row,
    "A" = "Bs",
    "B" = "Bs+NI",
    "C" = "NI",
    "D" = "Bs+Pp+NI",
    "E" = "Bs+Pp",
    "F" = "Pp+NI",
    "G" = "Pp",
    "H" = "M9"
  )
  
  # Final tidy
  plate_concat <- plate_concat[, .(well, datetime, condition, wavelength, value)]
  
  return(plate_concat)
}
```

# Apply to all files and concatenate

```{r}
#| output: false
#| label: 1da51b8d

files <- c(
  "008.0-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.1A-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.1B-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.2A-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.2B-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.3A-PP BS Growth single point_20260430_174232.xlsx",
  "008.3B-PP BS Growth single point_20260430_181943.xlsx",
  "008.4A-PP BS Growth single point_20260502_234619.xlsx",
  "008.4B-PP BS Growth single point_20260503_011632.xlsx"
)

all_data <- rbindlist(lapply(files, process_plate))

# elapsed time in hours (first value = 0)
all_data$time <- as.numeric(
  difftime(all_data$datetime, all_data$datetime[1], units = "hours") )
```

# Apply to only the "A" files and concatenate

```{r}
#| output: false
#| label: e5b861a6

files <- c(
  "008.0-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.1A-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.2A-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.3A-PP BS Growth single point_20260430_174232.xlsx",
  "008.4A-PP BS Growth single point_20260502_234619.xlsx"
)

all_A_data <- rbindlist(lapply(files, process_plate))

# elapsed time in hours (first value = 0)
all_A_data$time <- as.numeric(
  difftime(all_A_data$datetime, all_data$datetime[1], units = "hours") )
```

# Apply to only the "B" files and concatenate

```{r}
#| output: false

files <- c(
  "008.0-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.1B-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.2B-PP BS Growth single point (Modified)_20260424_160703.xlsx",
  "008.3B-PP BS Growth single point_20260430_181943.xlsx",
  "008.4B-PP BS Growth single point_20260503_011632.xlsx"
)

all_B_data <- rbindlist(lapply(files, process_plate))

# elapsed time in hours (first value = 0)
all_B_data$time <- as.numeric(
  difftime(all_B_data$datetime, all_B_data$datetime[1], units = "hours") )
```

# Write to CSV

```{r}
write.csv(all_data, "all_data.csv", row.names = FALSE)
#all_data <- read.csv ("all_data.csv") # Import CSV
```

# Summary statistics

```{r}
summary_df <- all_data %>% group_by(condition, wavelength, time) %>% summarise( mean_value = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE), n = n(), se = sd / sqrt(n), .groups = "drop" )
```

```{r}
summary_B_df <- all_B_data %>% group_by(condition, wavelength, time) %>% summarise( mean_value = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE), n = n(), se = sd / sqrt(n), .groups = "drop" )
```

```{r}
summary_A_df <- all_A_data %>% group_by(condition, wavelength, time) %>% summarise( mean_value = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE), n = n(), se = sd / sqrt(n), .groups = "drop" )
```

# Data Visualization

## Plot OD600 of all conditions BEFORE media replenishment

```{r}
#| echo: false
#| label: 008-plot_all_OD600_before
#| title: "OD<sub>600</sub> of all conditions BEFORE media replenishment"
summary_A_df %>% filter(wavelength %in% c("OD600")) %>% ggplot(aes(x = time, y = mean_value, color = condition)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  theme(title = element_markdown()) +
  labs(y="OD 600", x="time [h]", title = "OD<sub>600</sub> of all conditions BEFORE media replenishment")
```

## Plot OD600 of all conditions AFTER media replenishment

```{r}
#| echo: false
#| label: 008-plot_all_OD600_after

summary_B_df %>% filter(wavelength %in% c("OD600")) %>% ggplot(aes(x = time, y = mean_value, color = condition)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  theme(title = element_markdown()) +
  labs(y = expression("OD"[600]),
       x ="time [h]", 
       title = "OD<sub>600</sub> of all conditions AFTER media replenishment")
```

## Plot GFP of all conditions AFTER media replenishment

```{r}
#| echo: false
#| label: 008-plot_all_GFP70_after

summary_B_df %>% filter(wavelength %in% c("GFP70")) %>% ggplot(aes(x = time, y = mean_value, color = condition)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  theme(title = element_markdown()) +
  labs(y = "GFP",
       x ="time [h]", 
       title = "GFP intensity of all conditions AFTER media replenishment")
```

## Plot RFP of all conditions AFTER media replenishment

```{r}
#| echo: false
#| label: 008-plot_all_RFP_after

summary_B_df %>% filter(wavelength %in% c("OD600")) %>% ggplot(aes(x = time, y = mean_value, color = condition)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  theme(title = element_markdown()) +
  labs(y = "RFP",
       x ="time [h]", 
       title = "RFP of all conditions AFTER media replenishment")
```

## TO DO: 
- Substract M9 mean from values to normalize
- Calculate RFP/GFP per OD and compare Pp alone vs. Pp+NI to see how much NI is contributing to the OD






