---
title: "013-BPN_48h-replenish"
format: 
  html:
    code-copy: true
    code-fold: true
    code-summary: "Show code"
    output: false
    echo: false
execute:
    cache: true
itentifier: "013"
alias: "013"
lightbox: true
---

## Short description

Retry of experiment #008 with proper M9 media. In this experiment 7 different conditions — consisting of all possible combinations of *P. putida*, *B. subtilis* and a natural soil inoculum — were (co-)cultured in a well mixed environment (96-well-plate). Every 48 h the cultures were diluted 1:5 into fresh M9 media and further used as the new main culture. 12 biological replicates. After the fourth cycle cross contamination was observed.

## Data import and processing

Load dependencies.
```{r}
library(tidyverse)
library(readxl)
library(plotly)
library(ggnested)
library(ggtext)
#library(stringr)
#library(patchwork)
#library(RColorBrewer)
#library(growthcurver)
#library(data.table)
```

Process function for excel raw data from Tecan Spark plate reader (Toxic Lab):

```{r}
#| echo: true
#| code-fold: false

process_plate <- function(file_path) {
  raw <- read_excel(file_path, col_names = FALSE)
  
  # Extract plate sections
  OD600  <- raw[37:44, 2:13]
  GFP70 <- raw[70:77, 2:13]
  RFP85 <- raw[103:110, 2:13]
  RFP90  <- raw[136:143, 2:13]
  
  # Convert to long format
  to_long <- function(df, name) {
    out <- as.data.frame(as.table(as.matrix(df)))
    out$wavelength <- name
    return(out)
  }
  
  OD600  <- to_long(OD600,  "OD600")
  GFP70 <- to_long(GFP70, "GFP70")
  RFP85 <- to_long(RFP85, "RFP85")
  RFP90  <- to_long(RFP90,  "RFP90")
  
  # Combine
  plate_concat <- rbind(OD600, GFP70, RFP85, RFP90)
  plate_concat <- setNames(
    plate_concat,
    c("row", "col", "value", "wavelength")
    )
  
  plate_concat$value <- as.numeric(as.character(plate_concat$value))
  plate_concat$well <- paste0(plate_concat$row, plate_concat$col)
  
  # Date + time from same file
  date <- read_excel(file_path, range = "E6", col_names = FALSE)[[1]]
  time <- read_excel(file_path, range = "E7", col_names = FALSE)[[1]]
  
  plate_concat$datetime <- as.POSIXct(paste(date, time))
  
  # Conditions
  plate_concat$condition <- dplyr::recode(
    plate_concat$row,
    "A" = "B",
    "B" = "B+N",
    "C" = "N",
    "D" = "B+P+N",
    "E" = "B+P",
    "F" = "P+N",
    "G" = "P",
    "H" = "M9"
  )
  
  # Final tidy
  plate_concat <- plate_concat[, c("well", "datetime", "condition", "wavelength", "value")]
  
  return(plate_concat)
}
```

**Apply to all files and concatenate.**
```{r}
#| echo: true
#| code-fold: false

files <- list.files(pattern = "^013-.*\\.xlsx$", recursive = TRUE)

all_data <- dplyr::bind_rows(lapply(files, process_plate))


# all_data <- docall(rbind(lapply(files, process_plate)))

# elapsed time in hours (first value = 0)
all_data$time <- as.numeric(
  difftime(all_data$datetime, all_data$datetime[1], units = "hours") )
```
Apply to only the 'A' or 'B' files independently, to produce dataset before/after media replenishment for each timepoint.
```{r}

A_files <- list.files(pattern = "^013-.*\\.A.*\\.xlsx$", recursive = TRUE)
df_A <- dplyr::bind_rows(lapply(A_files, process_plate))

# elapsed time in hours (first value = 0)
df_A$time <- as.numeric(
  difftime(df_A$datetime, df_A$datetime[1], units = "hours") )


B_files <- list.files(pattern = "^013-.*\\.B.*\\.xlsx$", recursive = TRUE)
df_B <- dplyr::bind_rows(lapply(B_files, process_plate))

# elapsed time in hours (first value = 0)
df_B$time <- as.numeric(
  difftime(df_B$datetime, df_B$datetime[1], units = "hours") )
```
Write to `.csv`.
```{r}
write.csv(df_A, "df_before-dilution.csv", row.names = FALSE)
#all_data <- read.csv ("all_data.csv") # Import CSV
```
Calculate summary statistics.
```{r}
#| output: true
summary_df <- all_data %>% group_by(condition, wavelength, time) %>% summarise( mean_value = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE), n = n(), se = sd / sqrt(n), .groups = "drop" )

summary_B_df <- df_B %>%
  group_by(condition, wavelength, time) %>%
  summarise( mean_value = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE), n = n(), se = sd / sqrt(n), .groups = "drop" )

summary_A_df <- df_A %>% 
  group_by(condition, wavelength, time) %>% 
  summarise(
    mean_value = mean(value, na.rm = TRUE), 
    sd = sd(value, na.rm = TRUE), 
    n = n(), 
    se = sd / sqrt(n), 
    .groups = "drop"
    )
```

# Data Visualization
::: panel-tabset

## OD$_{600}$

```{r}
#| output: true
#| label: OD
#| fig-cap: OD$_{600}$ of all conditions before dilution/media replenishment.
#| lightbox:
#|   group: tab

OD <- summary_A_df %>% 
  filter(wavelength %in% c("OD600")) %>% 
  ggplot(aes(x = time, y = mean_value, color = condition)) +
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_minimal() +
  labs(y=expression("OD"[600]), x="time [h]")

OD
```

## GFP

```{r}
#| output: true
#| label: GFP
#| fig-cap: GFP of all conditions before media replenishment
#| lightbox:
#|   group: tab

GFP <- summary_A_df %>% 
  filter(wavelength %in% c("GFP70")) %>% 
  ggplot(aes(x = time, y = mean_value, color = condition)) +
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() +
  labs(y="GFP", x="time [h]")

GFP
```

## RFP
```{r}
#| output: true
#| label: RFP
#| fig-cap: RFP of all conditions before media replenishment
#| lightbox:
#|   group: tab

RFP <- summary_A_df %>% 
  filter(wavelength %in% c("RFP90")) %>% 
  ggplot(aes(x = time, y = mean_value, color = condition)) +
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() +
  labs(y="RFP", x="time [h]")

RFP
```
:::


```{r}
#| output: true
#| lightbox:
#|   group: all
OD
GFP
RFP
```


