---
title: "001-all_conditions_same_OD"
format: html
identifier: "001"
---

# Short Description

This is the first 96-well-plate experiment. All combinations of monoculture, and 2/3 species co-culture were tested with an initial starting OD600 of 0.001. Total well volume was 250 µl. Master mixes (MM) with OD of 0.005 were set up from overnight pre-culture and 50 µl MM used per well. 12 replicates (columns on plate) per condition. Order of conditions as follows:

|     |                 |
|-----|-----------------|
| A   | B.s.            |
| B   | B.s. + NI       |
| C   | NI              |
| D   | P.p + B.s. + NI |
| E   | P.p + B.s.      |
| F   | P.p + NI        |
| G   | P.p             |
| H   | control (M9)    |

: Rows of 96-well-plate

**Soil Preparation for NI** was done in H2O only. No Preculturing of natural inocoulum, no centrifugation. 1g soil (form TER lab course '26) + 10ml H2O were vortexted for 10 min and subsequently vacuum filtered through filter paper. Yield was very low. Cells were pelleted and resuspended in 2 ml M9 media.

# Define data loading and processing function
```{r}
library(readxl)
library(data.table)
library(dplyr)
library(ggplot2)
library(patchwork)
```

```{r}
process_plate <- function(file_path) {
  
  raw <- read_excel(file_path, col_names = FALSE)
  
  # Extract plate sections
  OD600  <- raw[25:32, 2:13]
  GFP100 <- raw[57:64, 2:13]
  GFP80  <- raw[89:96, 2:13]
  RFP100 <- raw[121:128, 2:13]
  RFP80  <- raw[153:160, 2:13]
  
  # Convert to long format
  to_long <- function(df, name) {
    out <- as.data.frame(as.table(as.matrix(df)))
    out$wavelength <- name
    return(out)
  }
  
  OD600  <- to_long(OD600,  "OD600")
  GFP100 <- to_long(GFP100, "GFP100")
  GFP80  <- to_long(GFP80,  "GFP80")
  RFP100 <- to_long(RFP100, "RFP100")
  RFP80  <- to_long(RFP80,  "RFP80")
  
  # Combine
  plate_concat <- rbindlist(list(OD600, GFP100, GFP80, RFP100, RFP80))
  setnames(plate_concat, c("row", "col", "value", "wavelength"))
  
  plate_concat$value <- as.numeric(as.character(plate_concat$value))
  plate_concat$well <- paste0(plate_concat$row, plate_concat$col)
  
  # Date + time from same file
  date <- read_excel(file_path, range = "B5", col_names = FALSE)[[1]]
  time <- read_excel(file_path, range = "B6", col_names = FALSE)[[1]]
  
  plate_concat$datetime <- strptime(
    paste(date, time),
    format = "%Y-%m-%d %I:%M:%S %p"
  )
  
  # Conditions
  plate_concat$condition <- dplyr::recode(
    plate_concat$row,
    "A" = "Bs", "B" = "Bs+Soil", "C" = "Soil",
    "D" = "Bs+Pp", "E" = "Bs+Pp+Soil",
    "F" = "Pp", "G" = "Pp+Soil", "H" = "M9"
  )
  
  # Final tidy
  plate_concat <- plate_concat[, .(well, datetime, condition, wavelength, value)]
  
  return(plate_concat)
}
```

# Apply to all files and concatenate

```{r}
#| echo: false

files <- c(
  "001-03-12-26_1605_BsPpSoil.xlsx",
  "001-03-12-26_1737_BsPpSoil.xlsx",
  "001-03-13-26_1133_BsPpSoil.xlsx",
  "001-03-13-26_1341_BsPpSoil.xlsx",
  "001-03-13-26_1530_BsPpSoil.xlsx",
  "001-03-13-26_1746_BsPpSoil.xlsx",
  "001-03-13-26_1930_BsPpSoil.xlsx",
  "001-03-14-26_1602_BsPpSoil.xlsx"
)

all_data <- rbindlist(lapply(files, process_plate))
```

# Add timepoint labels and elapsed time

```{r}
timepoints <- c("t0","t1.5","t19.5","t21.5","t23.5","t25.5","t27.5","t48")

all_data <- rbindlist(
  Map(function(f, tp) {
    dt <- process_plate(f)
    dt$timepoint <- tp
    dt
  }, files, timepoints)
)

# elapsed time in hours (first value = 0)
all_data$time <- as.numeric(
  difftime(all_data$datetime, all_data$datetime[1], units = "hours") )
```

# Write to CSV

```{r}
write.csv(all_data, "all_data.csv", row.names = FALSE)
#all_data <- read.csv ("all_data.csv") # Import CSV
```

# Summary statistics

```{r}
summary_df <- all_data %>% group_by(condition, wavelength, time) %>% summarise( mean_value = mean(value, na.rm = TRUE), sd = sd(value, na.rm = TRUE), n = n(), se = sd / sqrt(n), .groups = "drop" )
```

# Data Visualization

## Plot OD600

```{r}
summary_df %>% filter(wavelength %in% c("OD600")) %>% ggplot(aes(x = time, y = mean_value, color = condition)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  labs(y="OD 600", x="Time", title = "OD 600 of all conditions")
```

# OD600 for all Bacillus

```{r}
ggplot( summary_df %>% filter( condition %in% c("Bs", "Bs. + Soil", "Bs+Pp", "Bs+Pp+Soil") & wavelength %in% c("OD600") ), aes(x = time, y = mean_value, color = condition) ) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  labs(y = "OD 600", x = "Time", title = "OD 600 of Bacillus containing communities")
```

# OD600 for all Pseudomonas

```{r}
ggplot( summary_df %>% filter( condition %in% c("Pp", "Pp+Soil", "Bs+Pp", "Bs+Pp+Soil") & wavelength %in% c("OD600") ), aes(x = time, y = mean_value, color = condition) ) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  labs(y = "OD 600", x = "Time", title = "OD 600 of Pseudomonas containing communities")
```

# Plot GFP

```{r}
p3 <- summary_df %>% filter(wavelength %in% c("GFP100")) %>% 
  ggplot(aes(x = time, y = mean_value, color = condition)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  labs(y="GFP", x="Time") + theme(legend.position = "none")

# Plot RFP

p4 <- summary_df %>% filter(wavelength %in% c("RFP100")) %>% 
  ggplot(aes(x = time, y = mean_value, color = condition)) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  theme_classic() + 
  labs(y="RFP", x="Time")

p3 + p4 + plot_annotation(title="GFP and RFP intensities")
```

# Facet per condition

```{r}

pGFP <-summary_df %>% filter(wavelength == c("GFP100", "GFP80")) %>%
ggplot(aes(time, mean_value, color = wavelength)) + 
  geom_line() + 
  geom_point() +
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.2 ) +
  theme_classic() + 
  labs(y="GFP", x="Time") +
  facet_wrap(~ condition)

pRFP <-summary_df %>% filter(wavelength == c("RFP100", "RFP80")) %>%
ggplot(aes(time, mean_value, color = wavelength)) + 
  geom_line() + 
  geom_point() +
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.2 ) +
  theme_classic() +  
  labs(y="RFP", x="Time") +
  facet_wrap(~ condition)

pGFP + pRFP
```

# B. subtilis OD & RFP
```{r}
# (p1 | p2) / (p3 | p4) # For patchwork 2x2 grid
# Relative widths/heights can be adjusted:
# p1 + p2 + plot_layout(widths = c(2,1)) # makes first plot wider
#
# Overall annotation:
# (p1 | p2) / (p3 | p4) + plot_annotation(title="Fluorescence Plate Analysis")

p1 <- ggplot( summary_df %>% filter( condition %in% c("Bs", "Bs. + Soil", "Bs+Pp", "Bs+Pp+Soil") & wavelength %in% c("OD600") ), aes(x = time, y = mean_value, color = condition) ) + 
  geom_line() + geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  labs(y = "OD 600", x = "Time") + 
  theme_classic() +
  theme(legend.position = "none")

p2 <- summary_df %>% filter(condition %in% c("Bs", "Bs. + Soil", "Bs+Pp", "Bs+Pp+Soil") & wavelength %in% c("RFP100")) %>%
  ggplot(aes(x = time, y = mean_value, color = condition) ) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  labs(y = "RFP intensity", x = "Time") +
  theme_classic()

p1 + p2 + plot_annotation(title="OD 600 and RFP of Bacillus containing communities")
```

# P. putida OD & RFP

```{r}
p5 <- summary_df %>% filter(condition %in% c("Pp", "Pp+Soil", "Bs+Pp", "Bs+Pp+Soil") & wavelength %in% c("OD600")) %>%
  ggplot(aes(x = time, y = mean_value, color = condition) ) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) + 
  labs(y = "OD 600", x = "Time") + 
  theme(legend.position = "none") +
  theme_classic()

p6 <- summary_df %>% filter( condition %in% c("Pp", "Pp+Soil", "Bs+Pp", "Bs+Pp+Soil") & wavelength %in% c("GFP100") ) %>% 
  ggplot(aes(x = time, y = mean_value, color = condition) ) + 
  geom_line() + 
  geom_point() + 
  geom_errorbar( aes(ymin = mean_value - se, ymax = mean_value + se), width = 0.5 ) +
  labs(y = "GFP intensity", x = "Time") +
  theme_classic()

p5 + p6 + plot_annotation(title="OD 600 and RFP of Pseudomonas containing communities")
```



